---
title: "ggplot-Code-Along"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

Step 1:

```{r load packages}

# use the library function to load dplyr, tidyr, palmerpenguins, and ggplot2

library(tidyverse)


```

Steps 2 & 3:

```{r}
# Load the social Media dataset
# Dataset must be saved in the same place as your .Rmd to load data properly
# ../xxx/ tells relationally how files are related, .. go up file /into xxx/

socialMedia <- read_csv("../data/socialMediaAddiction.csv")

# Check out a "glimpse" of the variables of your dataframe

glimpse(socialMedia)
socialMedia

socialMediaMetadata <- read_csv("../data/socialMediaAddictionMetadata.txt")
socialMediaMetadata

```


Step 4:

```{r histogram}
#histograms for social media usage and sleep
#uses the tidyverse pipe %>% to connect statements, take this data and then do x
#when using gg plot and the tidyverse a + connects different functions
#aes is aesthetics, defines axis, colors, shapes
# add geometry data with geom_

socialMedia %>% 
  ggplot(mapping = aes(x = Avg_Daily_Usage_Hours)) +
  geom_histogram()

socialMedia %>% 
  ggplot(mapping = aes(x = Sleep_Hours_Per_Night)) +
  geom_histogram()
```
Step 5: The distributions of Avg_Daily_Usage_Hours and Sleep_Hours_Per_Night are both relatively normal. Avg_Daily_Usage_Hours has an extreme high count between 4-5 hours and Sleep_Hours_Per_Night has a very far fewer counts at the tails. I would hypothesize that Avg_Daily_Usage_Hours has a relatively normal distribution because . I would hypothesize that Sleep_Hours_Per_Night also has a relatively normal distribution because .

Step 6: I would create a scatterplot with social media usage on the x axis and hours of sleep on the y axis. This is because I want to see the correlation between these two variables and can include a line of best fit too.

Step 7:
```{r scatterplot}

# Scatterplot of bill length and weight
#geom smooth and geom smooth method for trendline

socialMedia %>%
  ggplot(mapping = aes(x = Sleep_Hours_Per_Night, y = Avg_Daily_Usage_Hours)) +
  geom_point() +
  geom_smooth(method = "lm")


# Scatterplot of bill length faceted by penguin species



```
Step 8: It appears that there is an indirect relationship between social media usage and sleep. The higher the social media usage, the less sleep an individual gets.

Step 9: I would use a box plot to visualize the association between social media app and usage time. I would have usage on the y axis and social media app on the x axis. This is because box plots can clearly demonstrate a quantitative variable divided according to categorical variable.
```{r boxplot}

# Plot boxplot of body mass by penguin Sex

socialMedia %>%
  ggplot(mapping = aes(x = Most_Used_Platform, y = Avg_Daily_Usage_Hours)) +
  geom_boxplot()

```
Step 11: Preferred social media app impacts social media addiction as individuals whose most used apps were WhatsApp had higher average usage than individuals whose most used apps were LINE or LinkedIn.

Step 12:

Step 13:

Step 14:
```{r}
# Plot boxplot of body mass by penguin species
#using color or fill equals x2 you can differentiate with a legend

penguins %>%
  ggplot(mapping = aes(x = species, y = body_mass_g, color = sex)) +
  geom_boxplot()  

# Plot boxplot of body mass by penguin Sex separated by species
#for multiple categorical variables
#facet has seperate plots for x1 with x2 plotted for each

penguins %>%
  ggplot(mapping = aes(x = species, y = body_mass_g)) +
  geom_boxplot() +
  facet_wrap(~sex)
```
Step 15: 